Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-10 vesicle_code/docs/adhesion/adhesion.tex
working on shear revisions

Figures~\ref{fig:doublet090-weakAdhesion}
and~\ref{fig:doublet090-strongAdhesion} show snapshots of two vesicles
that have formed a doublet and the color coding is the tension along the
vesicle.  The two vesicles are identical with reduced area $\Delta A =
0.9$, shear rate $\chi = 0.5$, separation distance $\delta = 0.4$, and
Hamaker constant $\mathcal{H} = 0.7$
(Figure~\ref{fig:doublet090-weakAdhesion}) and $\mathcal{H} = 2.1$
(Figure~\ref{fig:doublet090-strongAdhesion}).  The vesicles in the
doublet move in tandem rather than performing the tank-treading motion
that is observed in the absence of adhesion.  Similar to the quiescent
example, the membrane tension is negative in the contact region
indicating that the membrane is being compressed when the adhesive force
is strongest.

@@

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
separation distance to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In Figure~\ref{fig:sflow_distance}(a), we plot the
minimum distance between the vesicles as a function of time for various
Hamaker constants.  The red curves correspond to Hamaker constants that
are too small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.

In Figure~\ref{fig:sflow_distance}(b), we plot the time required for the
doublet to make a complete revolution at a variety of Hamaker constants
and shear rates.  For a fixed Hamaker constant, as expected, the period
decreases for larger shear rates.  In addition, for a fixed shear rate,
the period decreases as the Hamaker constant increases, but the period
eventually plateaus.  The dynamics of these doublet have similar
behavior to a single rigid body undergoing a Jeffery orbit in a shear
flow.  Therefore, we define the effective aspect ratio
$\lambda_{\mathrm{eff}}$ which satisfies
\begin{align*}
  \frac{2\pi}{\chi}\left(\lambda + \lambda^{-1} \right) = p,
\end{align*}
where $p$ is the period of the doublet.  The effective aspect ratio is
the aspect ratio of an ellipse in a shear flow that is undergoing a
Jeffery orbit with period $p$.  For the Hamaker constant $\mathcal{H} =
2$, the shear rates, periods, and effective aspect ratio are summarized
in Table~\ref{tbl:aspectRatio}.  Since the effective aspect ratio is
comparable for all three shear rates, this indicates that the dynamics
of a vesicle doublet in a shear flow resembles a rigid body undergoing a
Jeffery orbit.
\begin{table}[htp]
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 38.06 & 2.92 \\
  0.75 & 24.81 & 2.85 \\
  1.00 & 19.47 & 3.00 
\end{tabular}
\todo[inline]{Update these values with new periods.  Maybe need to
consider more than one Hamaker constant}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
Hamaker constant $\mathcal{H} = 2$, and separation distance $\delta =
0.4$.  The effective aspect ratios are comparable for all three shear
rates.}
\end{table}
\todo[inline]{period results with $\delta$ = 0.1, 0.2, 0.3, 0.4 so that
we can simply say how it affects the period}

>>>>

Figures~\ref{fig:doublet090-weakAdhesion}
and~\ref{fig:doublet090-strongAdhesion} show snapshots of two vesicles
that have formed a doublet and the color coding is the tension along the
vesicle.  Each of the vesicles have reduced area $\Delta A = 0.9$, shear
rate $\chi = 0.5$, separation distance $\delta = 0.4$, and Hamaker
constant $\mathcal{H} = 0.7$ (Figure~\ref{fig:doublet090-weakAdhesion})
and $\mathcal{H} = 2.1$ (Figure~\ref{fig:doublet090-strongAdhesion}).
Similar to the quiescent example, the membrane tension is negative in
the contact region indicating that the membrane is being compressed when
the adhesive force is strongest.  For both Hamaker constants, the
vesicles in the doublet move in tandem with a tank-reading motion that
is much slower than the tank-treading motion that occurs in the absense
of ahdesion.

@@

In Figure~\ref{fig:sflow_distance}(a), we plot the minimum distance
between the two vesicles as a function of time for various Hamaker
constants.  For Hamaker constants less than a critical value,
$\mathcal{H}_c$, the adhesive force is not strong enough to form a
doublet, and the distance between the vesicles grows with time (red
curves).  However, when $\mathcal{H} > \mathcal{H}_c$, the adhesive
force is sufficiently strong to bind the vesicles into a doublet, and
then the doublet undergoes a periodic motion (blue curves).  In
Figure~\ref{fig:sflow_distance}(b), we plot the time required for the
doublet to make a complete revolution for three different shear rates.
For a fixed Hamaker constant, as expected, the period decreases for
larger shear rates.  More interestingly, at a fixed shear rate, the
period depends non-monotonically on the Hamaker constant.  When the
Hamaker constant is first sufficiently large to form a doublet, the
individual vesicles undergo a sliding motion as they pass one another,
and the contact region between the vesicles moves along the vesicle
throughout the simulation (Figure~\ref{fig:doublet090-weakAdhesion}).
However, at larger Hamaker constants, once the contact region is formed,
the doublet maintains the same structure and undergoes a tumbling
dynamic similar to a Jeffery orbit.  The simulations illustrated in
Figures~\ref{fig:doublet090-weakAdhesion}
and~\ref{fig:doublet090-strongAdhesion} are chosen so that the doublet
has roughly the same period, but the structure of the doublet is
fundamentally different for these different Hamaker constants.

To further understand the behavior of the Jeffery orbit, we define an
effective aspect ratio $\lambda_{\mathrm{eff}}$ which satisfies
\begin{align*}
  \frac{2\pi}{\chi}\left(\lambda + \lambda^{-1} \right) = p,
\end{align*}
where $p$ is the period of the doublet.  The effective aspect ratio is
the aspect ratio of an ellipse in a shear flow that is undergoing a
Jeffery orbit with period $p$.  For the Hamaker constant $\mathcal{H} =
2$, the shear rates, periods, and effective aspect ratio are summarized
in Table~\ref{tbl:aspectRatio}.  Since the effective aspect ratio is
comparable for all three shear rates, this indicates that the dynamics
of a vesicle doublet in a shear flow resembles a rigid body undergoing a
Jeffery orbit.
\begin{table}[htp]
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 38.06 & 2.92 \\
  0.75 & 24.81 & 2.85 \\
  1.00 & 19.47 & 3.00 
\end{tabular}
\todo[inline]{Update these values with new periods.  Maybe need to
consider more than one Hamaker constant}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
Hamaker constant $\mathcal{H} = 2$, and separation distance $\delta =
0.4$.  The effective aspect ratios are comparable for all three shear
rates.}
\end{table}
\todo[inline]{period results with $\delta$ = 0.1, 0.2, 0.3, 0.4 so that
we can simply say how it affects the period}

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.
